---
title: 'Menatl Helath Cases in Karnataka Case Study '
output:
  pdf_document: default
  html_notebook: default
---

# Importing Data Set

```{r}

library(readr)

# Source : data.gov.in

DS1 <- read_csv("DataSet/District_Wise_Mental_Health_Patients_2018-19.csv")

head(DS1)
```

## 1. Top 10 District having a Suicide cases due to Mental Health Issues .

```{r}
library(ggplot2)

Sus <-data.frame("District"= DS1$DISTRICT,"Sucide_cases"=DS1$SUICIDE_ATTEMPT_CASES)


Sus_sort <- Sus[order(-Sus$Sucide_cases),]

top10 <- head(Sus_sort,10)

barplot(top10$Sucide_cases,names.arg = top10$District,xlab = "District",ylab = "Case Count")

```

## 2. Top 5 Districts having that possess high Mental Health Issues Cases due to Alcohol Consumption

```{r}
library(plotrix)

Alc <- data.frame("District" = DS1$DISTRICT,"No_of_cases"=DS1$`ALCOHOL_&_SUBSTANCE_ABUSE`) 

Alc_sort <- Alc[order(-Alc$No_of_cases),]

t10_alc <- head(Alc_sort,5)
par(cex = 2)
pie3D(t10_alc$No_of_cases,labels = t10_alc$District)
```

## 3. Construction of Boxplot to understand distribution of Common and Severe Mental Health Disorder.

```{r}

SMD <- DS1$`SEVERE_MENTAL_DISORDER_(SMD)`
CMD <- DS1 $`COMMON_MENTAL _DISORDER(CMD)`

boxplot(SMD)
boxplot(CMD)
```

| The first plot shows how **symmetrical** the **Common Mental Health Disorder** cases are distributed with Different cities
| The second plot shows how **symmetrical** the **Severe Mental Health Disorder** cases are distributed with Different cities .
| 

## 4. Performing linear regression factors affecting Severe Mental Disorder.

```{r}
model<- lm(DS1$`SEVERE_MENTAL_DISORDER_(SMD)`~DS1$`ALCOHOL_&_SUBSTANCE_ABUSE`)
summary(model)
```

| The model Concludes that the effect of Alcohol and Substance abuse effect is statistically significant.

The R-squared value of 0.6869 indicates that 68.69% of the variance in the dependent variable can be explained by the independent variable.

## 5.Performing Two tailed test to show the mean difference between Common Mental Health Disorder and Severe Mental Health Disorder.

```{r}

res <- t.test(SMD,CMD)
res
summary(res)

```

> The above test helps us to understand that there is a significant difference between Common mental Disorder and Severe Mental Health Disorder cases.
>
> The **p-value (0.03393)** is the probability of observing a t-statistic as extreme or more extreme than the one observed, assuming that the null hypothesis is true
